### 2014-10-07 paper/include/dtail.tex
Rewrite of DTail. Hopfully better!

Zehan Cui et al. \cite{dtail} propses DTail-R in which data for line refresh rate is stored in the DRAM itself instead of storing it in the memory controller which many other solutions do \cite{raidr}\cite{smartrefresh}\cite{refrint}. To counter the added latency of a slower memory they utilizes spatial locality of the refresh data and performs prefetching into a FIFO buffer. The retention time data consists of 4 bits and is interpreted in such way that the refresh period becomes \(64ms \times 2^n, n = [0...7]\). With this structure the amount of memory required is weighing in on a negliable amount of 0.006\% DRAM capacity (i.e. 1.92MB in a 32GB DRAM). One thing that DTail-R misses is refreshrate compensation for temperature variations. As the refresh rate is set per line DTail-R gets a higher resolution than techniques that set refreshrate on a higher level. Therefore DTail-R manages to decrease the amount of refreshes with as much as 87.9\% and thereby only have 16.7\% energy overhead compared to a refresh less 32GB DRAM. Zehan Cui et al. also proposes DTail-V \cite{dtail} which can be combined with DTail-R and is further discussed in \ref{sec:val}.

>>>>

Z. Cui et al. \cite{dtail} proposes DTail. They evaluates two versions of DTail referenced as DTail-R and DTail-V utilizing \textbf{R} and \textbf{V} data respecively. Even so, their main idea is that more types of refresh information can be incorperated and stored together with the \textbf{R} and \textbf{V} information. DTail can then choose the most efficient refresh schema depending on the current situation, i.e. workload and type of application. Refresh data is kept per row and is stored in the DRAM itself instead of storing it in the memory controller which many other solutions do \cite{raidr}\cite{smartrefresh}\cite{refrint}. To counter the added latency of a slower memory they utilizes spatial locality of the refresh data and performs prefetching into a FIFO buffer. As the refreash data is kept on a row level DTail can save more refreshes than other techniques that have a lower granularity.

For DTail-R, the retention time data consists of 3 bits and is interpreted in such way that the refresh period can become \(64ms \times 2^n, n = [0...7]\). With DTail-V one additional bit is added to the refresh data which is used to indicate if the row is valid or not. With this structure of four bits per row, the amount of memory required is weighing in on a negligible amount of 0.006\% DRAM capacity assuming a row size of 8 KB.

When a row is to be scheduled for refresh the refresh data is checked. If the row is valid and the time since the last refresh is \(64ms \times 2^n, n = [0...7]\) a refresh is performed, otherwise not. The refreshes can be done either by RAS commands or by auto refresh depending on what is most efficient at the time. If a large amount of rows needs refresh then auto refresh is selected, otherwise RAS-only refresh are used.

When only DTail-R is used, they manages to decrease the amount of refreshes with as much as 87.9\% and thereby only have 16.7\% energy overhead compared to a refresh less 32GB DRAM. With only DTail-V power consumption and performance overhead increase from 4\% and 3\% to 90\% and 23\% respectively when going from 10\% to 90\% memory capacity utilization compared to a refresh less 32 GB system. The power of DTail shows when DTail-R and DTail-V are used together and at 10\% memory utilization a performance overhead of only 1\% and close to 0\% increase in energy consumption is achived. When the utilization is increased to 90\% they increase to 2\% and 15\% respectively.
%

%One thing that DTail-R misses is refresh rate compensation for temperature variations.
